\begin{thebibliography}{1}
\bibitem{psl}
Broecheler, M., Mihalkova, L. \& Getoor, L.,
  Probabilistic Soft Logic.
In {\it Conference on Uncertainty in Artificial Intelligence}, 2010.

\bibitem{kuh}
 Kuh, G. D.,
What we're learning about student engagement from NSSE: Benchmarks for effective educational practices.
{\it Change: The Magazine of Higher Learning}, 2003.

\bibitem{carini}
Carini, R. M., Kuh, G. D. \& Klein, S. P.,
Student engagement and student learning: Testing the linkages.
{\it Research in Higher Education, 47(1), 1-32}, 2006.

\bibitem{kizilcec}
Kizilcec, R. F., Piech, C. \& Schneider, E.,
Deconstructing disengagement: analyzing learner subpopulations in massive open online courses.
In {\it International Conference on Learning Analytics and Knowledge}, 2013.

\bibitem{srl-book}
Getoor, L. \& Tasker, B.,
  Introduction to Statistical Relational Learning.
{\it The MIT Press}, 2007.

\bibitem{opinionfinder}
Wilson, T., Hoffmann, P., Somasundaran, S., Kessler, J., Wiebe, J., Choi, Y., Cardie, C., Riloff, E., \& Patwardhan, S.,
OpinionFinder: A System for Subjectivity Analysis.
In {\it Proceedings of
Human Language Technologies Conference/Conference on Empirical Methods
in Natural Language Processing (HLT/EMNLP 2005), Companion
	Volume (software demonstration)}, 2005.
	
\bibitem{bach-uai13}
Bach, S. H., Huang, B., London, B. \& Getoor, L.,
Hinge-loss Markov Random Fields: Convex Inference for Structured Prediction.
In {\it Uncertainty in Artificial Intelligence}, 2013.

\bibitem{bach-nips12}
Bach, S. H., Broecheler, M., Getoor, L. \& O'Leary, D. P.,
Scaling MPE Inference for Constrained Continuous Markov Random Fields.
In {\it Advances in Neural Information Processing Systems (NIPS) pages 2663--2671}, 2012.

\bibitem{bach-inferning13}
Bach, S. H., Huang, B. \& Getoor, L,
Learning Latent Groups with Hinge-loss Markov Random Fields.
In {\it Inferning: ICML Workshop on Interactions between Inference and Learning}, 2013.

\end{thebibliography} 